---
title: "R_Activity_Assignment_5"
author: "Dahee Ahn"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
```


## 1. Download the birdies data set from Canvas and load it into R using read.table(). We will be analyzing the weight gain of chicks (grams) as a function of different feeds
```{r}
birdies_df_txt <- read.table(file="C:/Users/chemk/OneDrive/Desktop/Classes/ENT6707_DataAnalysis/week6/birdies.txt", header=TRUE)
head(birdies_df_txt)
tail(birdies_df_txt)
str(birdies_df_txt)
nrow(birdies_df_txt)
View(birdies_df_txt)
birdies_df_txt$feed <- as.factor(birdies_df_txt$feed)
summary(birdies_df_txt)
str(birdies_df_txt)
```
```{r}
library(tidyverse)
birdies_df_txt %>% group_by(feed) %>% summarise(Means= mean(weight), SD=sd(weight), max_birdies=max(weight))
birdies_df_txt %>% group_by(feed) %>% summarise(Means= mean(weight), SD=sd(weight), max_birdies=max(weight)) %>% arrange(Means)
```

## 2. Fit an ANOVA of weight as a function of feed. Note that when I write blah1 as a function of blah2, blah1 should be the response variable and/or appear on the y-axis in any graphs whereas blah2 would be the predictor(s). Provide the R output for the ANOVA table and briefly explain how the degrees of freedom were calculated for each line of the table.
```{r}
windows(width=3, height=2)
library(datasets)
library(ggplot2)
head(birdies_df_txt)
plot(weight~feed, data=birdies_df_txt)
ggplot(birdies_df_txt, mapping=aes(x=feed, y=weight, fill=feed))+geom_boxplot()+theme_classic()+xlab("Feed")+ylab("Weight")
```

```{r}
hist(birdies_df_txt$weight)
qqnorm(birdies_df_txt$weight)
```
```{r}
var_check <- birdies_df_txt %>% group_by(feed) %>% summarize(Variance=var(weight))
var_check$Variance[2]/var_check$Variance[1]
```

```{r}
birdies_df_txt$sqrt_weight <- sqrt(birdies_df_txt$weight)
var_check_sqrt <- birdies_df_txt %>% group_by(feed) %>% summarize(Variance=var(sqrt_weight))
variance_ratio_sqrt <- var_check_sqrt$Variance[2]/var_check_sqrt$Variance[1]
print(variance_ratio_sqrt)
```

```{r}
fit_1 <- lm(weight~feed, data=birdies_df_txt)
anova(fit_1)
# Degrees of freedom of feed: k-1 (K = the number of groups) = 6-1 = 5
# Degrees of freedom of residauls: N-k (N = total number of observations) = 71-6 = 65
```
```{r}
plot(fit_1, which=c(2))
plot(fit_1, which=c(1))
```

## 3. The above ANOVA table you just created tells us if the variable weight varies across the levels of feed, but not HOW chick weight gain differs between feeds. Conduct pairwise comparisons using the emmeans package to identify any potential differences between groups (i.e., just report the R output for this question).
```{r}
library(emmeans)
emmeans(fit_1, pairwise~feed)
```

## 4. Write a BRIEF summary of your analysis by answering only the following: (i) which feed was associated with the largest chick weight gain? (ii) was it statistically clear that a single feed was the best for weight gain? Report the necessary statistics (t-values, degrees of freedom, and p-values) to justify your conclusions.

Q1. Which feed was associated with the largest chick weight gain?

A1. The feed associated with the largest chick weight gain was Sunflower, with an estimated mean weight of 329 gram.

Q2. Was it statistically clear that a single feed was the best for weight gain?

A2. There were some significant differences in weight gain among the different types of feed, while others were not. For instance: Comparing casein and horsebean, the weight(mean ± SE) of casein was 163.38g greater than that of horsebean (Tukey's test: t65 = 6.96,p < 0.0001), indicating a statistically significant difference. However, the weight(mean ± SE) of sunflower was only 5.33g greater than that of casein (Tukey's test: t65 = 0.24,p=0.9999), indicating no significant difference. This suggests that while casein perfomed significantly better than horsebean, there was no clear evidence that sunflower was statistically superior to casein in promoting weight gain. 

Q3. Report the necessary statistics (t-values, degrees of freedom, and p-values) to justify your conclusions.

A3. Comparing casein and horsebean: t-ratio: 6.96, df: 65, p-value: <0.0001 / Comparing casein and sunflower: t-ratio: 0.24, df: 65, p-value: 0.9999 / Comparing horsebean and sunflower: t-ratio: 7.18, df: 65, p-value: < 0.0001