---
title: "R_Activity_Assignment_3"
author: "Dahee Ahn"
date: "2024-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
```

```{r}
options(repos=c(CRAN="https://cran.r-project.org"))
install.packages("ggplot2")
```

## 1. Download the cherrytree data set from Canvas and load it into R using read.table(). Note this file is a .csv file (comma separated values).
```{r}
cherrytrees_df_csv <- read.table(file="C:/Users/chemk/OneDrive/Desktop/Classes/ENT6707_DataAnalysis/week4/cherrytrees.csv",header=TRUE, sep=",")
summary(cherrytrees_df_csv)
```
```{r}
head(cherrytrees_df_csv)
tail(cherrytrees_df_csv)
str(cherrytrees_df_csv)
```
## 2. Make a scatterplot of Height as a function of Girth using ggplot2. Use the theme theme_bw() and change the black circles to orange, hollow triangles.
```{r}
library("ggplot2")
ggplot(data=cherrytrees_df_csv, mapping=aes(x=Girth, y=Height))+geom_point(shape=2, col="orange")+theme_bw()
```
## 3.Make a boxplot of Height and as a function of Variety using ggplot2. Change the x-axis label to “Prunus variety” and the y-axis label to “Height (feet)”. Use the theme theme_classic() and fill the variety A box as dark blue and the variety B box as yellow.
```{r}
ggplot(cherrytrees_df_csv, mapping=aes(x=Variety, y=Height, fill=Variety))+geom_boxplot()+scale_fill_manual(values=c("A"="darkblue", "B"="yellow"))+theme_classic()+xlab("Prunus variety")+ylab("Height (feet)")
```


## 4. Make the same plot as the previous question, but make it a violin plot instead of a box plot and remove the legend.
```{r}
ggplot(cherrytrees_df_csv, mapping=aes(x=Variety, y=Height, fill=Variety))+geom_violin()+scale_fill_manual(values=c("A"="darkblue", "B"="yellow"))+theme_classic()+xlab("Prunus variety")+ylab("Height (feet)")+theme(legend.position="none")
```

