---
title: "R_Activity_Assignment_2"
author: "Dahee Ahn"
date: "2024-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
rm(list=ls())
```

## Question 1. Load in the breakfast data set. Remember to check the file type so that you can identify the right command for loading in the data.
```{r}
breakfast_df <- read.table("C:/Users/chemk/OneDrive/Desktop/Classes/ENT6707_DataAnalysis/week3/breakfast.txt", header = TRUE)
summary(breakfast_df)
nrow(breakfast_df)
head(breakfast_df)
tail(breakfast_df)
View(breakfast_df)
```

## Question 2. Use a single command to count the number of columns in the breakfast data.
```{r}
breakfast_df$Calories <- as.numeric(breakfast_df$Calories)
str(breakfast_df)
```

```{r}
ncol(breakfast_df)
```

## Question 3. Use base R (i.e., no external packages should be used for this question) to create a subset of the data that has only observations with values of Calories greater than 130.
```{r}
breakfast_df[which(breakfast_df$Calories>130),]
```

## Question 4. Use the pipe operators in tidyverse to create a subset of the data that has only observations with values of Calories greater than 130.
```{r}
library(tidyverse)
breakfast_df_subset_1 <- breakfast_df %>% filter(Calories > 130)
summary(breakfast_df_subset_1)
```

## Question 5. Use the pipe operators in tidyverse to calculate the mean Sugars by Company.
```{r}
breakfast_df %>% group_by(Company) %>% summarise(Means=mean(Sugars),) %>% arrange(Means)
```

